### 2024-09-24 starter_folder-main (1)/starter_folder-main/paper/paper.qmd
upload the starter files

---
title: "Demographic survey on Toronto shelter"
author: 
  - Xizi Sun
thanks: "A GitHub Repository containing all data, R code, and other files used in this investigation is located here: https://github.com/LilianS77/Toronto_shelter_research.git"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(opendatatoronto)
library(tidyverse)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| echo: false
#| warning: false
# Search for a specific package on OpenDataToronto by keyword ("Toronto Shelter System Flow")
trt_shelter <- search_packages("Toronto Shelter System Flow")
trt_shelter
# Retrieve a list of all resources (datasets) available within the found package
trt_shelter_resources <- trt_shelter %>%
  list_package_resources()
# Display the list of resources available in the "Toronto Shelter System Flow" package
trt_shelter_resources

trt_shelter <- trt_shelter_resources[2,] %>%
  get_resource()
data=trt_shelter

```

```{r}
#| echo: false
#| warning: false
#| label: fig-total-population
#| fig-cap: "Trend of Homeless Population in Toronto"
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Load the dataset (update with your file path)
data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/starter_folder-main (1)/data1.csv")

# Bar plot for 'returned_from_housing'
ggplot(data, aes(x = population_group, y = returned_from_housing, fill = population_group)) +
  geom_bar(stat = "identity") +
  labs(title = "Returned from Housing Across Population Groups",
       x = "Population Group", y = "Returned from Housing") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
# Bar plot for 'returned_to_shelter'
#| echo: false
#| warning: false
#| label: fig-total-population
#| fig-cap: "Trend of Homeless Population in Toronto"
ggplot(data, aes(x = population_group, y = returned_to_shelter, fill = population_group)) +
  geom_bar(stat = "identity") +
  labs(title = "Returned to Shelter Across Population Groups",
       x = "Population Group", y = "Returned to Shelter") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
#| echo: false
#| warning: false
#| label: fig-total-population
#| fig-cap: "Trend of Homeless Population in Toronto"
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Load the dataset (update the file path)
data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/starter_folder-main (1)/data1.csv")

# Convert the 'date.mmm.yy.' column to Date format
data$date <- as.Date(paste0("01-", data$date.mmm.yy.), format = "%d-%b-%y")

# Create a line plot to show the trend of a key population metric over time
ggplot(data, aes(x = date, y = actively_homeless, color = population_group, group = population_group)) +
  geom_line() +
  labs(title = "Trend of Actively Homeless Population Over Time",
       x = "Date", y = "Actively Homeless Count",
       color = "Population Group") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-09-24 starter_folder-main (1)/starter_folder-main/scripts/00-simulate_data.R

@@ -1,0 +1,42 @@
+#### Preamble ####
+# Purpose: Generates a simulated version of the Toronto Shelter
+# Author: Xizi Sun
+# Date: 23 September 2024 
+# Contact: xizi.sun@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Install the tidyverse
+# Any other information needed? no
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+
+#### Simulate data ####
+
+# Set seed for reproducibility
+
+set.seed(123)
+
+# Number of rows (change based on your dataset size)
+n <- 100  
+
+# Simulate "returned_from_housing" data: Assuming a normal distribution around 50 with some variation
+returned_from_housing <- round(rnorm(n, mean = 50, sd = 15))
+
+# Simulate "returned_to_shelter" data: Assuming a normal distribution around 100 with some variation
+returned_to_shelter <- round(rnorm(n, mean = 100, sd = 30))
+
+# Combine the simulated data into a data frame
+simulated_data <- data.frame(
+  returned_from_housing = returned_from_housing,
+  returned_to_shelter = returned_to_shelter
+)
+
+# View the first few rows of the simulated data
+head(simulated_data)
+
+# Optionally, save the simulated data to a CSV file
+write.csv(simulated_data, "/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/starter_folder-main (1)/simulated_data.csv")
+
+

### 2024-09-24 starter_folder-main (1)/starter_folder-main/scripts/01-download_data.R

@@ -1,0 +1,57 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from the City of Toronto’s open data portal.
+# Author: Xizi Sun
+# Date: 23 September 2024 
+# Contact: xizi.sun@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Install the tidyverse
+# Any other information needed? no
+
+
+#### Workspace setup ####
+library(opendatatoronto)
+library(tidyverse)
+
+
+#### Download data ####
+#We will use the `opendatatoronto` package to access a dataset from the City of Toronto’s open data portal.
+
+
+#| echo: false
+#| warning: false
+# Install necessary packages (if not installed)
+install.packages("opendatatoronto")
+# install.packages("tidyverse")
+library(opendatatoronto)
+library(tidyverse)
+
+# Retrieve a list of all available data packages from the OpenDataToronto portal
+packages <- list_packages(300)
+# Display the list of available packages
+packages
+
+
+#| echo: false
+#| warning: false
+# Search for a specific package on OpenDataToronto by keyword ("Toronto Shelter System Flow")
+trt_shelter <- search_packages("Toronto Shelter System Flow")
+trt_shelter
+# Retrieve a list of all resources (datasets) available within the found package
+trt_shelter_resources <- trt_shelter %>%
+  list_package_resources()
+# Display the list of resources available in the "Toronto Shelter System Flow" package
+trt_shelter_resources
+
+trt_shelter <- trt_shelter_resources[2,] %>%
+  get_resource()
+data=trt_shelter
+
+
+
+trt_shelter
+write.csv(trt_shelter,"/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/starter_folder-main (1)/trtdata1.csv")
+
+
+
+
+         

### 2024-09-24 starter_folder-main (1)/starter_folder-main/scripts/02-data_cleaning.R
upload the cleaned data and data test

@@ -1,44 +1,31 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw plane data recorded by 1 observer.
+# Author: Xizi Sun
+# Date: 23 September 2024 
+# Contact: xizi.sun@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Install the tidyverse
+# Any other information needed? no
+
 
 #### Workspace setup ####
 library(tidyverse)
 
-#### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+library(dplyr)
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+# Load the dataset (replace with the correct file path)
+data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/starter_folder-main (1)/trtdata1.csv")
+
+# Remove the columns: "X_id", "became_inactive", "actively_homeless"
[…]

### 2024-09-24 starter_folder-main (1)/starter_folder-main/scripts/03-test_data.R

@@ -1,15 +1,31 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: test a simulated version of the Toronto Shelter
+# Author: Xizi Sun
+# Date: 23 September 2024 
+# Contact: xizi.sun@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Install the tidyverse
+# Any other information needed? no
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Test data ####
+# Assuming the ID column is named 'X_id' or another column name
+# Replace 'ID_column_name' with the actual name of your ID column if it's different
+
+# Check if the minimum value in the ID column is at least 1
+min_value <- min(cleaned_data$X)
+
+# Print the result
+if(min_value >= 1) {
+  print("The minimum value in the ID column is at least 1.")
+} else {
+  print("Warning: The minimum value in the ID column is less than 1.")
+}
+
+# Alternatively, simply print the minimum value of the ID column
+min_value
+

### 2024-09-24 paper/paper.qmd
organized folders of the repo

---
title: "Demographic survey on Toronto shelter"
author: 
  - Xizi Sun
thanks: "A GitHub Repository containing all data, R code, and other files used in this investigation is located here: https://github.com/LilianS77/Toronto_shelter_research.git"
date: today
date-format: long
abstract: "This paper analyzes data from the City of Toronto's Shelter Management Information System (SMIS) to explore homelessness trends and shelter usage patterns. Key metrics such as returns to shelters after housing and the actively homeless population are examined across different demographic groups. The findings reveal that non-refugees and single adults face significant challenges in maintaining stable housing, with high rates of return to shelters. Additionally, the data shows an increase in homelessness post-2020, likely influenced by economic disruptions during the COVID-19 pandemic. The study highlights the need for targeted interventions to address housing stability and reduce homelessness in Toronto."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(opendatatoronto)
library(tidyverse)
```

# Introduction

Homelessness is a pressing issue in major cities worldwide, and Toronto is no exception. Recent observations suggest a growing number of homeless individuals on Toronto streets, raising concerns about the dynamics of the shelter system, the demographics of those affected, and their pathways into homelessness. To address these concerns, this study aims to analyze the flow of individuals into and out of Toronto’s shelter system, with a particular focus on understanding the origins and characteristics of the homeless population.By analyzing the patterns in shelter use, this study seeks to answer key questions such as: where are these individuals coming from, what are their demographic characteristics, and how frequently do they cycle through the shelter system? Understanding these factors is critical for identifying gaps in services and making data-driven policy recommendations to reduce homelessness in Toronto. The ultimate goal, in line with the City’s vision, is to contribute to efforts that make homelessness in Toronto a rare, brief, and non-recurring experience.

In this paper, I will analyze key indicators such as the number of people returning to shelters after being housed, the rate of newly identified homeless individuals, and demographic breakdowns such as age, gender, and population groups (youth, families, refugees, Indigenous individuals, etc.). This analysis will help illuminate the pathways into homelessness and provide evidence for policies aimed at mitigating the problem.

This paper is organized into several sections, each contributing to a comprehensive understanding of homelessness in Toronto, based on the provided shelter system dataset. The structure includes an exploration of the dataset, visualizations of key metrics, analysis of trends and patterns, and a discussion of the implications of these findings.

# Data {#sec-data}
## Overview
The dataset used in this study is sourced from the City of Toronto’s Shelter Management Information System (SMIS). This system tracks individuals who access the city's network of overnight shelters, respites, hotel/motel programs, and other allied services. The dataset provides monthly records on the movement of people into and out of these services, offering insight into the demographics, trends, and dynamics of homelessness in Toronto.

## Data Measurement
The dataset used in this study comes from the City of Toronto’s Shelter Management Information System (SMIS), which tracks individuals using shelters, respites, and other homeless services. The key variables and measurements include:

1. Shelter Flow Metrics
Returned from Housing: Measures the number of people who were housed but returned to shelters, indicating the challenges in maintaining stable housing.
Returned to Shelter: Tracks the number of people returning to shelters after leaving, providing insight into repeat shelter use.
Newly Identified: Captures individuals entering the shelter system for the first time, highlighting new cases of homelessness.
Moved to Housing: Counts people transitioning from shelters into permanent housing, a key measure of the system’s success in reducing homelessness.
2. Population Groups
Chronic Homelessness: Long-term homelessness, defined as 180 days in the past year or 546 days over three years.
Families, Youth, and Refugees: Specific groups are tracked to tailor interventions and support services based on different needs.
Indigenous Identity: Tracks those self-identifying as First Nations, Métis, or Inuit, given their disproportionate representation in the homeless population.
\newpage

# result

```{r}
#| echo: false
#| warning: false
#| label: fig-returned-from-housing
#| fig-cap: "Returned from Housing Across Population Groups"
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Load the dataset (update with your file path)
data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/starter_folder-main (1)/data1.csv")

# Bar plot for 'returned_from_housing'
ggplot(data, aes(x = population_group, y = returned_from_housing, fill = population_group)) +
  geom_bar(stat = "identity") +
  labs(title = "Returned from Housing Across Population Groups",
       x = "Population Group", y = "Returned from Housing") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
### figure 1
This bar plot shows the number of individuals from different population groups who returned to shelters after previously being housed. Key insights from the graph:

The All Population group has the highest number of returns from housing, with over 4,000 individuals.
Non-refugees and Single Adults also have a significant number of returns, indicating challenges for these groups in maintaining permanent housing.
Other population groups, such as Chronic, Families, and Youth, have much lower numbers of individuals returning from housing, but this might reflect their smaller total numbers in the population or different levels of support they receive.
```{r}
#| echo: false
#| warning: false
#| label: fig-returned-to-shelter
#| fig-cap: "Returned to Shelter Across Population Groups"
ggplot(data, aes(x = population_group, y = returned_to_shelter, fill = population_group)) +
  geom_bar(stat = "identity") +
  labs(title = "Returned to Shelter Across Population Groups",
       x = "Population Group", y = "Returned to Shelter") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
### firgure 2
This bar plot visualizes the number of individuals returning to shelters across various population groups. Key insights:

The All Population group dominates again, with over 30,000 people returning to shelters.
Non-refugees and Single Adults show similarly high returns, indicating that these groups have more difficulty staying housed.
Chronic, Indigenous, and Families groups have fewer returns, although these groups may face other challenges in accessing shelter or permanent housing.
Youth shows relatively low numbers of returns to shelters.
```{r}
#| echo: false
#| warning: false
#| label: fig-total-population
#| fig-cap: "Trend of Homeless Population in Toronto"
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Load the dataset (update the file path)
data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/starter_folder-main (1)/data1.csv")

# Convert the 'date.mmm.yy.' column to Date format
data$date <- as.Date(paste0("01-", data$date.mmm.yy.), format = "%d-%b-%y")

# Create a line plot to show the trend of a key population metric over time
ggplot(data, aes(x = date, y = actively_homeless, color = population_group, group = population_group)) +
  geom_line() +
  labs(title = "Trend of Actively Homeless Population Over Time",
       x = "Date", y = "Actively Homeless Count",
       color = "Population Group") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
### firgure 3
This line plot shows the trends in the number of actively homeless individuals across different population groups from 2018 to 2024. Key trends:

The All Population group consistently has the highest number of actively homeless individuals over time, though there are fluctuations in the count.
Non-refugees and Single Adults have seen a significant increase in actively homeless individuals since 2020, which may indicate worsening conditions or increased difficulty in accessing housing.
Chronic and Families groups have more stable trends, while Youth and Indigenous groups show more modest and stable numbers.
The trend highlights the overall increase in homelessness post-2020 for several groups, which could be linked to broader economic or social factors affecting housing availability.

# Discussion
This analysis sheds light on the patterns and dynamics of homelessness in Toronto, drawing on data from the City’s shelter system. The key findings indicate significant challenges for certain population groups in maintaining permanent housing, frequent returns to shelters, and a concerning rise in homelessness over time, particularly after 2020.

## First discussion point {#sec-first-point}
The Returned from Housing analysis revealed that individuals from the Non-refugee and Single Adult groups face persistent barriers in sustaining housing stability, with high return rates to shelters. This suggests that these groups may not be receiving sufficient support or resources, such as access to affordable housing, stable employment opportunities, or necessary mental health services. Additionally, the overall high return rate across the All Population group reflects the broader systemic issue of housing insecurity in Toronto, where many individuals struggle to transition from temporary shelters to permanent housing.

## Second discussion point
The Returned to Shelter plot reinforced these findings, highlighting how certain population groups—particularly Non-refugees and Single Adults—are cycling in and out of the shelter system more frequently. This revolving-door effect suggests that while shelters provide temporary relief, they are not effectively breaking the cycle of homelessness for many individuals. The lower return rates for Chronic, Families, and Youth might indicate more successful interventions for these groups, though their continued presence in shelters points to a need for further long-term solutions

## Third discussion point
The Trend of Actively Homeless Population Over Time provides a broader context for understanding how homelessness in Toronto has evolved. The significant rise in homelessness after 2020, especially among Non-refugees and Single Adults, may be attributed to the economic impact of the COVID-19 pandemic, including job losses, eviction increases, and reduced access to social services. This trend underscores the vulnerability of certain groups during times of economic instability and highlights the need for responsive, proactive policies that address housing affordability, economic security, and social support.

\newpage

\appendix

# Appendix {-}
## Dataset and Graph Sketches {#sec-appendix-sketches}

Sketches depicting both the desired dataset and the graphs generated in this analysis are available in the GitHub Repository.

## Data Cleaning {#sec-appendix-cleaning}

The data cleaning process involved filtering out some of the columns from the raw dataset and renaming some of the data entries for clarity and simplicity.

# Additional data details

\newpage

# References

### 2024-09-24 paper/paper.qmd
update paper writing

@@
 # Load the dataset (update with your file path)
-data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/starter_folder-main (1)/data1.csv")
+data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/data/analysis_data/trtdata1.csv")
 
@@
 # Load the dataset (update the file path)
-data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/starter_folder-main (1)/data1.csv")
+data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/data/analysis_data/trtdata1.csv")
 

### 2024-09-24 scripts/00-simulate_data.R
update simulate data

@@ -37,6 +37,6 @@ simulated_data <- data.frame(
 head(simulated_data)
 
 # Optionally, save the simulated data to a CSV file
-write.csv(simulated_data, "/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/starter_folder-main (1)/simulated_data.csv")
+write.csv(simulated_data, "/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/data/analysis_data/simulated_data.csv")
 
 

### 2024-09-26 scripts/00-simulate_data.R
download the raw data and simulate data

@@ -7,36 +7,40 @@
 # Pre-requisites: Install the tidyverse
 # Any other information needed? no
 
+# Load necessary libraries
+library(dplyr)
 
-#### Workspace setup ####
-library(tidyverse)
+# Define the population groups
+population_groups <- c("All Population", "Chronic", "Refugees", "Families", "Youth", "Single Adult", "Non-refugees")
 
-
-#### Simulate data ####
-
-# Set seed for reproducibility
-
+# Set up the simulation for 12 months (2018), monthly data
 set.seed(123)
+n_months <- 12  # Simulate for 12 months
+dates <- format(seq(as.Date("2018-01-01"), by = "month", length.out = n_months), "%b-%y")
 
-# Number of rows (change based on your dataset size)
-n <- 100  
-
-# Simulate "returned_from_housing" data: Assuming a normal distribution around 50 with some variation
-returned_from_housing <- round(rnorm(n, mean = 50, sd = 15))
-
-# Simulate "returned_to_shelter" data: Assuming a normal distribution around 100 with some variation
-returned_to_shelter <- round(rnorm(n, mean = 100, sd = 30))
-
-# Combine the simulated data into a data frame
+# Create simulated data without the removed columns
 simulated_data <- data.frame(
-  returned_from_housing = returned_from_housing,
-  returned_to_shelter = returned_to_shelter
+  date.mmm = rep(dates, each = length(population_groups)),
+  population_group = rep(population_groups, times = n_months),
+  newly_identified = sample(0:1200, n_months * length(population_groups), replace = TRUE),
+  moved_to_housing = sample(0:600, n_months * length(population_groups), replace = TRUE),
+  age_under_16 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
+  age_16_24 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
+  age_25_34 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
+  age_35_44 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
+  age_45_54 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
+  age_55_64 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
+  age_65over = sample(0:1500, n_months * length(population_groups), replace = TRUE)
 )
+
+# Preview the simulated data
+head(simulated_data)
+
+# Optionally, save to a CSV
[…]

### 2024-09-26 scripts/01-download_data.R

@@ -49,7 +49,7 @@ data=trt_shelter
 
 
 trt_shelter
-write.csv(trt_shelter,"/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/starter_folder-main (1)/trtdata1.csv")
+write.csv(trt_shelter,"/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/data/raw_data/raw_data.csv")
 
 
 

### 2024-09-27 scripts/02-data_cleaning.R
update cleaned data

@@ -8,24 +8,19 @@
 # Any other information needed? no
 
 
-#### Workspace setup ####
-library(tidyverse)
-
+# Load necessary library
 library(dplyr)
 
-# Load the dataset (replace with the correct file path)
-data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/starter_folder-main (1)/trtdata1.csv")
+# Load the raw data (replace 'raw_data.csv' with your actual file path if different)
+raw_data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/data/raw_data/raw_data.csv")
 
-# Remove the columns: "X_id", "became_inactive", "actively_homeless"
-cleaned_data <- data %>%
-  select(-X_id, -became_inactive, -actively_homeless)
+# Perform data cleaning: keep only the columns that were simulated
+cleaned_data <- raw_data %>%
+  select(date.mmm.yy., population_group, newly_identified, moved_to_housing, 
+         ageunder16, age16.24, age25.34, age35.44, age45.54, age55.64, age65over)
 
-# Display the first few rows of the cleaned data
+# Preview the cleaned data
 head(cleaned_data)
 
-# Optionally, write the cleaned data to a new CSV file
-write.csv(cleaned_data, "cleaned_trtdata1.csv", row.names = FALSE)
-
-
 #### Save data ####
-write_csv(cleaned_data, "/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/starter_folder-main (1)/starter_folder-main/data/analysis_data/cleaned_data.csv")
+write_csv(cleaned_data, "/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/data/analysis_data/analysis_data.csv")

### 2024-09-27 scripts/00-simulate_data.R
updated simulated data

@@ -13,24 +13,24 @@ library(dplyr)
 # Define the population groups
 population_groups <- c("All Population", "Chronic", "Refugees", "Families", "Youth", "Single Adult", "Non-refugees")
 
-# Set up the simulation for 12 months (2018), monthly data
+# Set up the simulation for 12 months (2018), monthly data with the new date format 'date.mmm.yy'
 set.seed(123)
 n_months <- 12  # Simulate for 12 months
 dates <- format(seq(as.Date("2018-01-01"), by = "month", length.out = n_months), "%b-%y")
 
-# Create simulated data without the removed columns
+# Create simulated data
 simulated_data <- data.frame(
-  date.mmm = rep(dates, each = length(population_groups)),
+  date.mmm.yy = rep(dates, each = length(population_groups)),
   population_group = rep(population_groups, times = n_months),
   newly_identified = sample(0:1200, n_months * length(population_groups), replace = TRUE),
   moved_to_housing = sample(0:600, n_months * length(population_groups), replace = TRUE),
-  age_under_16 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
-  age_16_24 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
-  age_25_34 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
-  age_35_44 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
-  age_45_54 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
-  age_55_64 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
-  age_65over = sample(0:1500, n_months * length(population_groups), replace = TRUE)
+  ageunder16 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
+  age16.24 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
+  age25.34 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
+  age35.44 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
+  age45.54 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
+  age55.64 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
+  age65over = sample(0:1500, n_months * length(population_groups), replace = TRUE)
 )
 
 # Preview the simulated data

### 2024-09-27 scripts/03-test_data.R
update test data

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: test a simulated version of the Toronto Shelter
+# Purpose: test both simulated and cleaned data
 # Author: Xizi Sun
 # Date: 23 September 2024 
 # Contact: xizi.sun@mail.utoronto.ca
@@ -8,24 +8,92 @@
 # Any other information needed? no
 
 
-#### Workspace setup ####
-library(tidyverse)
+# Load necessary libraries
+library(dplyr)
 
+### Part 1: Test the Simulated Data
 
-#### Test data ####
-# Assuming the ID column is named 'X_id' or another column name
-# Replace 'ID_column_name' with the actual name of your ID column if it's different
+# Load the simulated data
+simulated_data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/data/raw_data/simulated_data.csv")
 
-# Check if the minimum value in the ID column is at least 1
-min_value <- min(cleaned_data$X)
+cat("----Testing Simulated Data----\n")
 
-# Print the result
-if(min_value >= 1) {
-  print("The minimum value in the ID column is at least 1.")
+# 1. Test for duplicate rows
+if (any(duplicated(simulated_data))) {
+  cat("There are duplicate rows in the simulated data.\n")
 } else {
-  print("Warning: The minimum value in the ID column is less than 1.")
+  cat("No duplicate rows in the simulated data.\n")
 }
 
-# Alternatively, simply print the minimum value of the ID column
-min_value
+# 2. Test for missing values (partial missingness)
+missing_values_simulated <- colSums(is.na(simulated_data))
+cat("Missing values per column in simulated data:\n")
+print(missing_values_simulated)
 
+# 3. Test the range of the date column
+simulated_data$date.mmm.yy <- as.Date(paste0("01-", simulated_data$date.mmm.yy), format = "%d-%b-%y")
+date_range_simulated <- range(simulated_data$date.mmm.yy)
+cat("Date range in simulated data is from:", date_range_simulated[1], "to", date_range_simulated[2], "\n")
+
+# 4. Test the range of 'newly_identified' and 'moved_to_housing'
+cat("Range of newly_identified in simulated data:", range(simulated_data$newly_identified), "\n")
+cat("Range of moved_to_housing in simulated data:", range(simulated_data$moved_to_housing), "\n")
+
[…]

### 2024-09-27 scripts/00-simulate_data.R
update simulated data

@@ -22,8 +22,8 @@ dates <- format(seq(as.Date("2018-01-01"), by = "month", length.out = n_months),
 simulated_data <- data.frame(
   date.mmm.yy = rep(dates, each = length(population_groups)),
   population_group = rep(population_groups, times = n_months),
-  newly_identified = sample(0:1200, n_months * length(population_groups), replace = TRUE),
-  moved_to_housing = sample(0:600, n_months * length(population_groups), replace = TRUE),
+  returned_to_shelter = sample(0:1200, n_months * length(population_groups), replace = TRUE),
+  returned_from_housing = sample(0:600, n_months * length(population_groups), replace = TRUE),
   ageunder16 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
   age16.24 = sample(0:1500, n_months * length(population_groups), replace = TRUE),
   age25.34 = sample(0:1500, n_months * length(population_groups), replace = TRUE),

### 2024-09-27 scripts/02-data_cleaning.R
update cleaned data

@@ -16,7 +16,7 @@ raw_data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/d
 
 # Perform data cleaning: keep only the columns that were simulated
 cleaned_data <- raw_data %>%
-  select(date.mmm.yy., population_group, newly_identified, moved_to_housing, 
+  select(date.mmm.yy., population_group, returned_from_housing, returned_to_shelter, 
          ageunder16, age16.24, age25.34, age35.44, age45.54, age55.64, age65over)
 
 # Preview the cleaned data

### 2024-09-27 scripts/03-test_data.R
test both simulated data and cleaned data

@@ -35,15 +35,15 @@ simulated_data$date.mmm.yy <- as.Date(paste0("01-", simulated_data$date.mmm.yy),
 date_range_simulated <- range(simulated_data$date.mmm.yy)
 cat("Date range in simulated data is from:", date_range_simulated[1], "to", date_range_simulated[2], "\n")
 
-# 4. Test the range of 'newly_identified' and 'moved_to_housing'
-cat("Range of newly_identified in simulated data:", range(simulated_data$newly_identified), "\n")
-cat("Range of moved_to_housing in simulated data:", range(simulated_data$moved_to_housing), "\n")
+# 4. Test the range of 'returned_to_shelter' and 'returned_from_housing'
+cat("Range of returned_to_shelter in simulated data:", range(simulated_data$returned_to_shelter), "\n")
+cat("Range of returned_from_housing in simulated data:", range(simulated_data$returned_from_housing), "\n")
 
-# 5. Check for outliers in 'newly_identified' and 'moved_to_housing'
-outliers_newly_identified <- simulated_data$newly_identified[simulated_data$newly_identified > (mean(simulated_data$newly_identified) + 3*sd(simulated_data$newly_identified))]
-outliers_moved_to_housing <- simulated_data$moved_to_housing[simulated_data$moved_to_housing > (mean(simulated_data$moved_to_housing) + 3*sd(simulated_data$moved_to_housing))]
-cat("Outliers in 'newly_identified':", outliers_newly_identified, "\n")
-cat("Outliers in 'moved_to_housing':", outliers_moved_to_housing, "\n")
+# 5. Check for outliers in 'returned_to_shelter' and 'returned_from_housing'
+outliers_returned_to_shelter <- simulated_data$returned_to_shelter[simulated_data$returned_to_shelter > (mean(simulated_data$returned_to_shelter) + 3*sd(simulated_data$returned_to_shelter))]
+outliers_returned_from_housing <- simulated_data$returned_from_housing[simulated_data$returned_from_housing > (mean(simulated_data$returned_from_housing) + 3*sd(simulated_data$returned_from_housing))]
+cat("Outliers in 'returned_to_shelter':", outliers_returned_to_shelter, "\n")
+cat("Outliers in 'returned_from_housing':", outliers_returned_from_housing, "\n")
 
 # 6. Summary statistics for numeric columns
[…]

### 2024-09-27 scripts/00-simulate_data.R
update new simulated data

@@ -19,6 +19,7 @@ n_months <- 12  # Simulate for 12 months
 dates <- format(seq(as.Date("2018-01-01"), by = "month", length.out = n_months), "%b-%y")
 
 # Create simulated data
+# Generated by ChatGPT
 simulated_data <- data.frame(
   date.mmm.yy = rep(dates, each = length(population_groups)),
   population_group = rep(population_groups, times = n_months),
@@ -36,8 +37,8 @@ simulated_data <- data.frame(
 # Preview the simulated data
 head(simulated_data)
 
-# Optionally, save to a CSV
-write.csv(simulated_data,"/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/data/raw_data/simulated_data.csv")
+#  save to a CSV
+write_csv(simulated_data, here::here("data", "raw_data", "simulated_data.csv"))
 
 # View the first few rows of the simulated data
 head(simulated_data)

### 2024-09-27 scripts/01-download_data.R
update the scripts to download the dataset

@@ -4,7 +4,7 @@
 # Date: 23 September 2024 
 # Contact: xizi.sun@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Install the tidyverse
+# Pre-requisites: Install the opendatatoronto (Gelfand 2022) and tidyverse (Wickham et al. 2019) packages.
 # Any other information needed? no
 
 
@@ -49,7 +49,7 @@ data=trt_shelter
 
 
 trt_shelter
-write.csv(trt_shelter,"/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/data/raw_data/raw_data.csv")
+write_csv(trt_shelter, here::here("data", "raw_data", "raw_data.csv"))
 
 
 

### 2024-09-27 scripts/02-data_cleaning.R
update the scripts to do data cleaning

@@ -4,15 +4,15 @@
 # Date: 23 September 2024 
 # Contact: xizi.sun@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Install the tidyverse
+# Pre-requisites: Run the file "01-download_data.R".
 # Any other information needed? no
 
 
 # Load necessary library
 library(dplyr)
 
-# Load the raw data (replace 'raw_data.csv' with your actual file path if different)
-raw_data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/data/raw_data/raw_data.csv")
+# Load the raw data 
+raw_data <- read_csv(here::here("data", "raw_data", "raw_data.csv")) %>%
 
 # Perform data cleaning: keep only the columns that were simulated
 cleaned_data <- raw_data %>%
@@ -23,4 +23,4 @@ cleaned_data <- raw_data %>%
 head(cleaned_data)
 
 #### Save data ####
-write_csv(cleaned_data, "/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/data/analysis_data/analysis_data.csv")
+write_csv(cleaned_data, here::here("data", "analysis_data", "analysis_data.csv"))

### 2024-09-27 scripts/03-test_data.R
update test data

@@ -4,7 +4,7 @@
 # Date: 23 September 2024 
 # Contact: xizi.sun@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Install the tidyverse
+# Pre-requisites: Run the files "01-download_data.R" and "02-data_cleaning.R".
 # Any other information needed? no
 
 
@@ -14,7 +14,8 @@ library(dplyr)
 ### Part 1: Test the Simulated Data
 
 # Load the simulated data
-simulated_data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/data/raw_data/simulated_data.csv")
+simulated_data <- read_csv(here::here("data", "raw_data", "simulated_data.csv"))
+
 
 cat("----Testing Simulated Data----\n")
 
@@ -58,8 +59,7 @@ print(unique_population_groups)
 ### Part 2: Test the Cleaned Data
 
 # Load the cleaned data
-cleaned_data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/data/analysis_data/analysis_data.csv")
-
+cleaned_data <- read_csv(here::here("data", "analysis_data", "analysis_data.csv"))
 cat("\n----Testing Cleaned Data----\n")
 
 # 1. Test for duplicate rows

### 2024-09-27 paper/paper.qmd
update part of paper

@@
 date-format: long
-abstract: "This paper analyzes data from the City of Toronto's Shelter Management Information System (SMIS) to explore homelessness trends and shelter usage patterns.
-Key metrics such as returns to shelters after housing and the actively homeless population are examined across different demographic groups.
-The findings reveal that non-refugees and single adults face significant challenges in maintaining stable housing, with high rates of return to shelters.
-Additionally, the data shows an increase in homelessness post-2020, likely influenced by economic disruptions during the COVID-19 pandemic.
-The study highlights the need for targeted interventions to address housing stability and reduce homelessness in Toronto."
+abstract: "重写·"
 format: pdf
@@
 
-library(tidyverse)
-library(palmerpenguins)
 library(opendatatoronto)
 library(tidyverse)
+library(knitr)
 ```
 
-# Introduction
+# Introduction {#sec-introduction}
 
@@
 # Data {#sec-data}
-## Overview
-The dataset used in this study is sourced from the City of Toronto’s Shelter Management Information System (SMIS).
-This system tracks individuals who access the city's network of overnight shelters, respites, hotel/motel programs, and other allied services.
-The dataset provides monthly records on the movement of people into and out of these services, offering insight into the demographics, trends, and dynamics of homelessness in Toronto.
 
-## Data Measurement
-The dataset used in this study comes from the City of Toronto’s Shelter Management Information System (SMIS), which tracks individuals using shelters, respites, and other homeless services.
-The key variables and measurements include:
+## Data Tools {#sec-data-tools}
+All data analysis and visualizations were conducted using R [@citeR].
+Data was obtained from the City of Toronto's Open Data Catalogue [@opendata] and included the Toronto Shelter System Flow dataset [@torontosheltersystem].
+Visualizations were created using the ggplot2 package [@ggplot2], with data manipulation carried out via dplyr [@dplyr] and tidyr [@tidyr].
+Date handling was simplified with the lubridate package [@lubridate], and all R code was formatted using the styler package [@styler].
+Guidance on storytelling with data was drawn from Telling Stories with Data [@tellingstorieswithdata].
 
-1. Shelter Flow Metrics
-Returned from Housing: Measures the number of people who were housed but returned to shelters, indicating the challenges in maintaining stable housing.
-Returned to Shelter: Tracks the number of people returning to shelters after leaving, providing insight into repeat shelter use.
-Newly Identified: Captures individuals entering the shelter system for the first time, highlighting new cases of homelessness.
-Moved to Housing: Counts people transitioning from shelters into permanent housing, a key measure of the system’s success in reducing homelessness.
-2. Population Groups
-Chronic Homelessness: Long-term homelessness, defined as 180 days in the past year or 546 days over three years.
-Families, Youth, and Refugees: Specific groups are tracked to tailor interventions and support services based on different needs.
-Indigenous Identity: Tracks those self-identifying as First Nations, Métis, or Inuit, given their disproportionate representation in the homeless population.
+## Data Source And Measurement {#sec-data-source-and-measurement}
+The dataset Toronto Shelter System Flow[@torontosheltersystem] used in this study is sourced from the Toronto Open Data Platform[@opendata].
+It measures and tracks homelessness in Toronto by capturing information through a Shelter Management Information System (SMIS) that covers various services such as shelters, 24-hour respite sites, and warming centers funded by the City of Toronto.
+The data spans from January 2018 to September 2024 and includes individuals who have accessed any of these services in the past three months.
+The dataset updated monthly and published on the 15th of each month, provides monthly records on the movement of people into and out of these services, offering insight into the demographics, trends, and dynamics of homelessness in Toronto.
+
+The table [@tbl-cleaned-data] shows the variables after data cleaning, showing 6 of the 604 observations.
+Due to space constraints we are only showing 2 age groups in this table.
+The variable “date.mmm.yyy.” goes by month from 2018 to August 2024.
+The data covers different age groups: under 16, 16-24, 25-34, 35-44, 45-54, 55-64, 65 and over, and types of population: refugees, single adults, families, youth, and chronically ill.
+It tracks inflow and outflow: returned from shelter and returned from housing.
+
+```{r}
+#| label: tbl-cleaned-data
+#| tbl-cap: Sample of cleaned lead data 
+#| echo: false
+
+# Load the dataset
+cleaned_data <- read_csv(here::here("data", "analysis_data", "analysis_data.csv"), show_col_types = FALSE)
+
+# Display the first 6 rows of selected columns and match with column names
+head(cleaned_data[, c("date.mmm.yy.", "population_group", "returned_from_housing", "returned_to_shelter", "ageunder16", "age16.24")], 6) |>
+  kable(
+    col.names = c("Date", "Population Group", "Returned from Housing", "Returned to Shelter", "Age under 16", "Age 16-24"),
+    booktabs = TRUE
+  )
+
+```
+
+## Data Selection
+
+There are many databases on homelessness in Toronto, such as About Hostel Services: Homeless Shelter Locations[@toronto2019hostelservices], but it doesn't distinguish well between population types, and we can't tell from it where homeless people come from.
+The data [@torontosheltersystem] I chose has very complete age groupings and also divides the people who seek help from shelters like they return from housing or return to shelter.
+better still we can know more information about the homeless from this dataset such as whether they are refugees, single or have a family so that we can combine it with the literature to have a better discussion about the Homeless trends and causes of homelessness in Toronto.
+
 \newpage
@@
 #| warning: false
-#| label: fig-returned-from-housing
-#| fig-cap: "Returned from Housing Across Population Groups"
-# Load necessary libraries
+#| label: fig-total-homeless-trend-by-quarter
+#| fig-cap: "Total Homeless Population Trend by Quarter"
 library(ggplot2)
 library(dplyr)
+library(lubridate)
 
-# Load the dataset (update with your file path)
-data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/data/analysis_data/trtdata1.csv")
+# Generated by ChatGPT
 
-# Bar plot for 'returned_from_housing'
-ggplot(data, aes(x = population_group, y = returned_from_housing, fill = population_group)) +
-  geom_bar(stat = "identity") +
-  labs(title = "Returned from Housing Across Population Groups",
-       x = "Population Group", y = "Returned from Housing") +
+# Define the age group columns
+age_columns <- c("ageunder16", "age16.24", "age25.34", "age35.44", "age45.54", "age55.64", "age65over")
+
+# Convert the date column to a proper date format, extract year and quarter, and filter out 2024 Q3
+total_homeless_quarter_trend <- cleaned_data %>%
+  mutate(
+    date = as.Date(paste0("01-", date.mmm.yy.), format = "%d-%b-%y"),
+    year = year(date),
+    quarter = quarter(date)
+  ) %>%
+  filter(!(year == 2024 & quarter == 3)) %>%
+  group_by(year, quarter) %>%
+  summarise(total_homeless = sum(across(all_of(age_columns)), na.rm = TRUE), .groups = "drop") %>%
+  mutate(period = paste0(year, " Q", quarter)) # Create a combined year-quarter label
+
+# Create a line plot to show the total homeless population trend by quarter (excluding 2024 Q3)
+ggplot(total_homeless_quarter_trend, aes(x = period, y = total_homeless, group = 1)) +
+  geom_line(color = "blue") +
+  geom_point(color = "blue") +
+  labs(
+    x = "Year-Quarter", y = "Total Homeless Population"
+  ) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
 
-```
-### figure 1
-This bar plot shows the number of individuals from different population groups who returned to shelters after previously being housed. Key insights from the graph:
-
-The All Population group has the highest number of returns from housing, with over 4,000 individuals.
-Non-refugees and Single Adults also have a significant number of returns, indicating challenges for these groups in maintaining permanent housing.
-Other population groups, such as Chronic, Families, and Youth, have much lower numbers of individuals returning from housing, but this might reflect their smaller total numbers in the population or different levels of support they receive.
+@fig-total-homeless-trend-by-quarter shows the total homeless population trend from 2018 Q1 to 2024 Q2.
+The population rises until 2019, then sharply declines in 2020, with the lowest point falling below 60,000.
+From 2021 onward, the population steadily increases, reaching its highest point in 2024 Q2, exceeding 80,000.
 ```{r}
@@
 #| fig-cap: "Returned to Shelter Across Population Groups"
-ggplot(data, aes(x = population_group, y = returned_to_shelter, fill = population_group)) +
+
+# Load dplyr to ensure filter works
+library(dplyr)
+
+#Generated by ChatGPT
+
+# Assuming 'cleaned_data' or 'data' is your dataframe
+ggplot(
+  cleaned_data %>% dplyr::filter(population_group != "All Population"),
+  aes(x = population_group, y = returned_to_shelter, fill = population_group)
+) +
   geom_bar(stat = "identity") +
-  labs(title = "Returned to Shelter Across Population Groups",
-       x = "Population Group", y = "Returned to Shelter") +
+  labs(
+  x = "Population Group", y = "Returned to Shelter"
+  ) +
   theme_minimal() +
@@
 ```
+@fig-returned-to-shelter 
+```{r}
+#| echo: false
+#| warning: false
+#| label: fig-returned-from-housing
+#| fig-cap: "Returned from Housing Across Population Groups"
+
+# Load dplyr to ensure the correct filter function is used
+library(dplyr)
+
+# Ensure you are referencing the correct dataset, like 'cleaned_data'
+ggplot(
+  cleaned_data %>% dplyr::filter(population_group != "All Population"),
+  aes(x = population_group, y = returned_from_housing, fill = population_group)
+) +
+  geom_bar(stat = "identity") +
+  labs(
+    title = "Returned from Housing Across Population Groups",
+    x = "Population Group", y = "Returned from Housing"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
 ### firgure 2
-This bar plot visualizes the number of individuals returning to shelters across various population groups. Key insights:
+visualizes the number of individuals returning to shelters across various population groups. Key insights:
 
@@
 Youth shows relatively low numbers of returns to shelters.
+
 ```{r}
@@
 #| warning: false
-#| label: fig-total-population
-#| fig-cap: "Trend of Homeless Population in Toronto"
-# Load necessary libraries
+#| label: fig-age-group-distribution-single-adult-pie
+#| fig-cap: "Age Group Distribution in Single Adult Population"
 library(ggplot2)
 library(dplyr)
+library(lubridate)
 
-# Load the dataset (update the file path)
-data <- read.csv("/Users/XiziS/OneDrive/Desktop/1/Toronto_shelter_research/data/analysis_data/trtdata1.csv")
+#generated by ChatGPT
 
-# Convert the 'date.mmm.yy.' column to Date format
-data$date <- as.Date(paste0("01-", data$date.mmm.yy.), format = "%d-%b-%y")
+# Define the age group columns
+age_columns <- c("ageunder16", "age16.24", "age25.34", "age35.44", "age45.54", "age55.64", "age65over")
 
-# Create a line plot to show the trend of a key population metric over time
-ggplot(data, aes(x = date, y = actively_homeless, color = population_group, group = population_group)) +
-  geom_line() +
-  labs(title = "Trend of Actively Homeless Population Over Time",
-       x = "Date", y = "Actively Homeless Count",
-       color = "Population Group") +
+# Filter for Single Adult Population, adjust this filter condition based on your actual dataset structure
+single_adult_population <- cleaned_data %>%
+  filter(population_group == "Single Adult") %>%
+  pivot_longer(cols = all_of(age_columns), names_to = "Age Group", values_to = "Count") %>%
+  group_by(`Age Group`) %>%
+  summarise(total_population = sum(Count, na.rm = TRUE), .groups = "drop")
+
+# Calculate percentages
+single_adult_population <- single_adult_population %>%
+  mutate(percentage = total_population / sum(total_population) * 100,
+         label = paste0(round(percentage, 1), "%"))  # Create a label with percentage
+
+# Plot the pie chart
+ggplot(single_adult_population, aes(x = "", y = percentage, fill = `Age Group`)) +
+  geom_bar(stat = "identity", width = 1) +
+  coord_polar(theta = "y") +
+  labs(
+    title = "Age Group Distribution in Single Adult Population"
+  ) +
+  geom_text(aes(label = label), position = position_stack(vjust = 0.5)) +  # Add percentage labels
+  theme_void() +  # Remove unnecessary axes
+  theme(legend.position = "right") +
+  scale_fill_brewer(palette = "Set3")  # Optional: Add color palette
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| label: fig-age-group-proportion-trend-yearly
+#| fig-cap: "Trends in Age Group Proportions Over Time (Yearly)"
+library(ggplot2)
+library(dplyr)
+library(lubridate)
+
+#developed by ChatGPT
+
+# Define the age group columns
+age_columns <- c("ageunder16", "age16.24", "age25.34", "age35.44", "age45.54", "age55.64", "age65over")
+
+# Convert the date column to a proper date format, extract year, and calculate the total population and proportions for each age group
+age_group_proportion_trend_yearly <- cleaned_data %>%
+  mutate(
+    date = as.Date(paste0("01-", date.mmm.yy.), format = "%d-%b-%y"),
+    year = year(date)
+  ) %>%
+  pivot_longer(cols = all_of(age_columns), names_to = "Age Group", values_to = "Count") %>%
+  group_by(year, `Age Group`) %>%
+  summarise(total_population = sum(Count, na.rm = TRUE), .groups = "drop") %>%
+  group_by(year) %>%
+  mutate(Proportion = total_population / sum(total_population)) %>%
+  ungroup()
+
+# Create a line plot to show the age group proportion trend over time (yearly)
+ggplot(age_group_proportion_trend_yearly, aes(x = factor(year), y = Proportion, color = `Age Group`, group = `Age Group`)) +
+  geom_line(size = 1.2) +
+  geom_point() +
+  scale_y_continuous(labels = scales::percent) +  # Format y-axis as percentages
+  labs(
+    x = "Year", y = "Proportion of Population"
+  ) +
   theme_minimal() +
@@
 ```
-### firgure 3
-This line plot shows the trends in the number of actively homeless individuals across different population groups from 2018 to 2024. Key trends:
+
+```
+ shows the trends in the number of actively homeless individuals across different population groups from 2018 to 2024. Key trends:
 

### 2024-09-27 paper/paper.qmd
update paper draft

@@
 ```
-@fig-returned-to-shelter 
+@fig-returned-to-shelter describes a significantly higher number of non-refugees and single adults returning to shelter compared to other groups.
+Approximately 30,000 non-refugees returned to shelter, the highest value in the chart.
+Single adults are slightly below 30,000, but still make up a large percentage of the population returning to shelters.
+Refugee, family, chronically ill, indigenous and youth groups returned to shelters in significantly lower numbers.
 ```{r}
@@
   labs(
-    title = "Returned from Housing Across Population Groups",
     x = "Population Group", y = "Returned from Housing"
@@
 ```
-### firgure 2
-visualizes the number of individuals returning to shelters across various population groups. Key insights:
-
-The All Population group dominates again, with over 30,000 people returning to shelters.
-Non-refugees and Single Adults show similarly high returns, indicating that these groups have more difficulty staying housed.
-Chronic, Indigenous, and Families groups have fewer returns, although these groups may face other challenges in accessing shelter or permanent housing.
-Youth shows relatively low numbers of returns to shelters.
+@fig-returned-from-housing shows shelter returns among different demographic groups, and similar to returns to shelters, non-refugees had the highest number of returns from shelter, just over 4,000.
+The number of single adults is also high at just under 4,000.
+Chronically ill and family populations are moderately represented, with both returning from shelters in the range of 1,000 to 1,500 people.
+The indigenous and youth groups have the lowest rates of return from housing, with just over 500 indigenous people and just over 1,000 youth.
 
@@
 ```
-
+@fig-age-group-distribution-single-adult-pie shows the distribution of age groups within the current single adult homeless population, highlighting the predominance of the 25.34 and 35.44 year old age groups, which account for approximately 25.2% and 25.8% of the total population respectively.
+In contrast, the 16.24 year old age group is relatively underrepresented at 8.5%.
+65 years and over are not represented in any of the single adult populations, accounting for 0%.
+The under-16s are minors and therefore account for 0%.
+This distribution suggests that middle-aged people form the bulk of the population, while the young and the old are virtually absent.
 ```{r}
@@
 ```
+@fig-age-group-proportion-trend-yearly shows several important trends in the homeless population by age group.
+The 25-34 and 35-44 age groups have been consistently higher than other age groups since 2018 and are growing, especially after 2021.
+25-44 year olds make up a growing share of the homeless population, totaling more than 50 percent of the total in 2024.
+At the same time, the number of homeless people in the middle-aged age group of 45.54 to 55.64 years is decreasing, and the 65+ age group is trending downwards after 2022.
+Children and young people (under 16) will start to decline after 2018 but will increase after 2021.
 
-```
- shows the trends in the number of actively homeless individuals across different population groups from 2018 to 2024. Key trends:
+\newpage
 
-The All Population group consistently has the highest number of actively homeless individuals over time, though there are fluctuations in the count.
-Non-refugees and Single Adults have seen a significant increase in actively homeless individuals since 2020, which may indicate worsening conditions or increased difficulty in accessing housing.
-Chronic and Families groups have more stable trends, while Youth and Indigenous groups show more modest and stable numbers.
-The trend highlights the overall increase in homelessness post-2020 for several groups, which could be linked to broader economic or social factors affecting housing availability.
+# Discussion {#sec-data}
 
-# Discussion
-This analysis sheds light on the patterns and dynamics of homelessness in Toronto, drawing on data from the City’s shelter system.
-The key findings indicate significant challenges for certain population groups in maintaining permanent housing, frequent returns to shelters, and a concerning rise in homelessness over time, particularly after 2020.
+## Overview of Homeless in Toronto  {#sec-overview-of-Toronto-homeless}
+Rising rates of homelessness in Toronto reflect significant challenges in managing housing instability, particularly among specific vulnerable groups.
+According to a comprehensive literature review on homelessness in Toronto, the number of homeless people has increased dramatically over the past few decades [@gaetz2004homelessness].
+The data @fig-returned-to-shelter and @fig-returned-from-housing shows that non-refugees and single adults make up the largest percentage of those returning to shelters, demonstrating their vulnerability and continued housing insecurity.
+This trend is likely caused by economic challenges, lack of affordable housing, and inadequate social safety nets.
+The literature [@gaetz2004homelessness] highlights how policy changes, including cuts to social housing and benefits, have exacerbated the problem, putting more people at risk of chronic homelessness.
 
-## First discussion point {#sec-first-point}
-The Returned from Housing analysis revealed that individuals from the Non-refugee and Single Adult groups face persistent barriers in sustaining housing stability, with high return rates to shelters.
-This suggests that these groups may not be receiving sufficient support or resources, such as access to affordable housing, stable employment opportunities, or necessary mental health services.
-Additionally, the overall high return rate across the All Population group reflects the broader systemic issue of housing insecurity in Toronto, where many individuals struggle to transition from temporary shelters to permanent housing.
+## Impact of unemployment on single adults aged 25-44 years {#sec-Impact-of-Unemployment}
+In @fig-returned-to-shelter and @fig-returned-from-housing, we see that single adults make up a large portion of those returning to shelters for help.
+And from @fig-age-group-distribution-single-adult-pie and @fig-age-group-proportion-trend-yearly, 25-44 year olds make up a large portion of single adults and the overall homeless population.
+The unemployment crisis has severely impacted single adults, especially those aged 25-44, who make up a large portion of the homeless population.The COVID-19 pandemic triggered a sharp economic downturn that led to widespread unemployment and income insecurity.
+According to [@jones2020canadianlabourmarket], Canada's unemployment rate rose sharply in 2020, with young and middle-aged adults disproportionately affected.
+This population is particularly vulnerable because they rely heavily on employment to pay rent and meet basic needs.
+Without a steady income, many in this age group face eviction, bankruptcy, and an increased likelihood of homelessness.
+This is compounded by mental health challenges and limited access to social services [@poremski2015employment].
 
-## Second discussion point
-The Returned to Shelter plot reinforced these findings, highlighting how certain population groups—particularly Non-refugees and Single Adults—are cycling in and out of the shelter system more frequently.
-This revolving-door effect suggests that while shelters provide temporary relief, they are not effectively breaking the cycle of homelessness for many individuals.
-The lower return rates for Chronic, Families, and Youth might indicate more successful interventions for these groups, though their continued presence in shelters points to a need for further long-term solutions
-
-## Third discussion point
-The Trend of Actively Homeless Population Over Time provides a broader context for understanding how homelessness in Toronto has evolved.
-The significant rise in homelessness after 2020, especially among Non-refugees and Single Adults, may be attributed to the economic impact of the COVID-19 pandemic, including job losses, eviction increases, and reduced access to social services.
-This trend underscores the vulnerability of certain groups during times of economic instability and highlights the need for responsive, proactive policies that address housing affordability, economic security, and social support.
-
+## Future Insights {#sec-Future-Insights}
+Given the rising rates of homelessness among single adults, public policy interventions must address both immediate and long-term solutions.
+Expanding affordable housing and strengthening employment support programs are critical.
+Governments at all levels should also focus on improving access to mental health services and implementing policies that prevent evictions, especially in times of economic crises.
+A particular emphasis should be placed on preventative measures, such as emergency rental assistance programs and increasing the availability of affordable housing [@gaetz2004homelessness].
+Additionally, the development of job retraining programs tailored to individuals who have lost employment due to economic downturns could provide sustainable solutions for those in the 25-44 age bracket.
+## Weakness and Limitations {#sec-Weakness-and-Limitations}
+While the data provides valuable insights into homelessness trends in Toronto, there are important limitations that should be acknowledged.
+The dataset reflects only individuals who use overnight shelter services, excluding those who sleep outdoors or rely on non-shelter-based services.
+Additionally, shelters that do not participate in Toronto's Shelter Management Information System (SMIS) are not included in the dataset.
+As a result, the data may underestimate the total homeless population, especially among individuals who do not seek help through formal shelter systems.
 \newpage
@@
 
-# Additional data details
+# Statement on LLM usage {#sec-LLMs}
+
+Aspects of the code were written with the help of the auto-complete tool, ChatGPT.
+The abstract, introduction and discusstion were written with the help of ChatGPT and the entire chat history is available in other/llms/usage.txt.
 
@@
 
-# References
+# References {#sec-references}

### 2024-09-27 paper/paper.qmd
update paper

@@
 ---
-title: "Demographic survey on Toronto shelter"
+title: "Who's Homeless in Toronto: Major Trends in Single Adults Aged 25 to 44 Becoming Homeless from 2018 to 2024"
 author: 
@@
 date-format: long
-abstract: "重写·"
+abstract: "This study examines homelessness trends in Toronto from 2018 to 2024, focusing on demographic shifts and socio-economic factors.
+Findings show a sharp rise in homelessness among single adults aged 25-44, largely due to post-pandemic unemployment.
+The research provides insights into how Toronto can better allocate resources to reduce homelessness by targeting vulnerable groups.
+Stronger public policies focused on expanding support services and employment opportunities for at-risk populations are recommended."
 format: pdf
+toc: true
 number-sections: true
@@
 library(knitr)
+library(ggplot2)
+library(dplyr)
+library(lubridate)
 ```
@@
 
-Homelessness is a pressing issue in major cities worldwide, and Toronto is no exception.
-Recent observations suggest a growing number of homeless individuals on Toronto streets, raising concerns about the dynamics of the shelter system, the demographics of those affected, and their pathways into homelessness.
-To address these concerns, this study aims to analyze the flow of individuals into and out of Toronto’s shelter system, with a particular focus on understanding the origins and characteristics of the homeless population.By analyzing the patterns in shelter use, this study seeks to answer key questions such as: where are these individuals coming from, what are their demographic characteristics, and how frequently do they cycle through the shelter system?
-Understanding these factors is critical for identifying gaps in services and making data-driven policy recommendations to reduce homelessness in Toronto.
-The ultimate goal, in line with the City’s vision, is to contribute to efforts that make homelessness in Toronto a rare, brief, and non-recurring experience.
+In recent years, the growing number of homeless people in Toronto has raised concerns and alarms, generating a high level of public concern and calls for action.
+This paper uses dataset to analyze age group trends and demographic dynamics of Toronto's homeless population, using the “Toronto Shelter System Flows” dataset [@torontosheltersystem] that tracks homelessness from 2018 to 2024.
+By focusing on how different age groups and populations are affected, the study delves into the rising rates of homelessness among single adults aged 25-44.
+It also discusses the post-pandemic rise in unemployment as a socio-economic factor contributing to these trends [@jones2020canadianlabourmarket].
 
-In this paper, I will analyze key indicators such as the number of people returning to shelters after being housed, the rate of newly identified homeless individuals, and demographic breakdowns such as age, gender, and population groups (youth, families, refugees, Indigenous individuals, etc.).
-This analysis will help illuminate the pathways into homelessness and provide evidence for policies aimed at mitigating the problem.
+Despite the growing awareness of Toronto's homelessness crisis, most existing research has not adequately discussed the changing composition of the homeless population.
+This study aims to fill this gap by analyzing these demographic changes in detail and linking them to broader societal trends.
+By identifying the groups most at risk, the study provides important insights into how the city of Toronto can better allocate resources to prevent homelessness.
 
-This paper is organized into several sections, each contributing to a comprehensive understanding of homelessness in Toronto, based on the provided shelter system dataset.
-The structure includes an exploration of the dataset, visualizations of key metrics, analysis of trends and patterns, and a discussion of the implications of these findings.
+In this report, @sec-data-tools describes the data tools I used to analyze the data, and @sec-data-source-and-measurement explains the context and measurements used with this dataset.
+@sec-data-selection explains the reasons for not choosing another database.
+@sec-Results presents and describes the demographic distribution and trends of homelessness in graphs and charts.
+@sec-Discussion analyzes the information in the graphs in the context of the community and describes the limitations of the study.
 
@@
 
-## Data Selection
+## Data Selection {#sec-data-selection}
 
@@
 
-\newpage
-
-# result
+# Result {#sec-result}
 
@@
 #| fig-cap: "Total Homeless Population Trend by Quarter"
-library(ggplot2)
-library(dplyr)
-library(lubridate)
 
@@
 ```
-
 @fig-total-homeless-trend-by-quarter shows the total homeless population trend from 2018 Q1 to 2024 Q2.
-The population rises until 2019, then sharply declines in 2020, with the lowest point falling below 60,000.
-From 2021 onward, the population steadily increases, reaching its highest point in 2024 Q2, exceeding 80,000.
+The population rises until 2019, then sharply declines in 2020, with the lowest point falling below 80,000.
+From 2021 onward, the population steadily increases, reaching its highest point in 2024 Q2, exceeding 120,000.
 ```{r}
@@
 
-# Load dplyr to ensure the correct filter function is used
-library(dplyr)
+#Generated by ChatGPT
 
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
 ```
@@
 #| fig-cap: "Age Group Distribution in Single Adult Population"
-library(ggplot2)
-library(dplyr)
-library(lubridate)
 
@@
   coord_polar(theta = "y") +
-  labs(
-    title = "Age Group Distribution in Single Adult Population"
-  ) +
   geom_text(aes(label = label), position = position_stack(vjust = 0.5)) +  # Add percentage labels
@@
 #| fig-cap: "Trends in Age Group Proportions Over Time (Yearly)"
-library(ggplot2)
-library(dplyr)
-library(lubridate)
 
@@
 Additionally, the development of job retraining programs tailored to individuals who have lost employment due to economic downturns could provide sustainable solutions for those in the 25-44 age bracket.
+
 ## Weakness and Limitations {#sec-Weakness-and-Limitations}
@@
 
-\appendix
-
-# Appendix {-}
-## Dataset and Graph Sketches {#sec-appendix-sketches}
-
-Sketches depicting both the desired dataset and the graphs generated in this analysis are available in the GitHub Repository.
-
-## Data Cleaning {#sec-appendix-cleaning}
-
-The data cleaning process involved filtering out some of the columns from the raw dataset and renaming some of the data entries for clarity and simplicity.
-
 # Statement on LLM usage {#sec-LLMs}
@@
 
-\newpage
-
 # References {#sec-references}

### 2024-09-27 scripts/00-simulate_data.R
styler scripts for simulated data

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Generates a simulated version of the Toronto Shelter
 # Author: Xizi Sun
-# Date: 23 September 2024 
+# Date: 23 September 2024
 # Contact: xizi.sun@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Install the tidyverse
@@ -15,7 +15,7 @@ population_groups <- c("All Population", "Chronic", "Refugees", "Families", "You
 
 # Set up the simulation for 12 months (2018), monthly data with the new date format 'date.mmm.yy'
 set.seed(123)
-n_months <- 12  # Simulate for 12 months
+n_months <- 12 # Simulate for 12 months
 dates <- format(seq(as.Date("2018-01-01"), by = "month", length.out = n_months), "%b-%y")
 
 # Create simulated data
@@ -42,6 +42,3 @@ write_csv(simulated_data, here::here("data", "raw_data", "simulated_data.csv"))
 
 # View the first few rows of the simulated data
 head(simulated_data)
-
-
-

### 2024-09-27 scripts/01-download_data.R
styler scripts for download_data

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from the City of Toronto’s open data portal.
 # Author: Xizi Sun
-# Date: 23 September 2024 
+# Date: 23 September 2024
 # Contact: xizi.sun@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Install the opendatatoronto (Gelfand 2022) and tidyverse (Wickham et al. 2019) packages.
@@ -14,7 +14,7 @@ library(tidyverse)
 
 
 #### Download data ####
-#We will use the `opendatatoronto` package to access a dataset from the City of Toronto’s open data portal.
+# We will use the `opendatatoronto` package to access a dataset from the City of Toronto’s open data portal.
 
 
 #| echo: false
@@ -42,16 +42,11 @@ trt_shelter_resources <- trt_shelter %>%
 # Display the list of resources available in the "Toronto Shelter System Flow" package
 trt_shelter_resources
 
-trt_shelter <- trt_shelter_resources[2,] %>%
+trt_shelter <- trt_shelter_resources[2, ] %>%
   get_resource()
-data=trt_shelter
+data <- trt_shelter
 
 
 
 trt_shelter
 write_csv(trt_shelter, here::here("data", "raw_data", "raw_data.csv"))
-
-
-
-
-         

### 2024-09-27 scripts/02-data_cleaning.R
styler scripts for data cleaning

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans the raw plane data recorded by 1 observer.
 # Author: Xizi Sun
-# Date: 23 September 2024 
+# Date: 23 September 2024
 # Contact: xizi.sun@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Run the file "01-download_data.R".
@@ -11,13 +11,14 @@
 # Load necessary library
 library(dplyr)
 
-# Load the raw data 
+# Load the raw data
 raw_data <- read_csv(here::here("data", "raw_data", "raw_data.csv")) %>%
-
-# Perform data cleaning: keep only the columns that were simulated
-cleaned_data <- raw_data %>%
-  select(date.mmm.yy., population_group, returned_from_housing, returned_to_shelter, 
-         ageunder16, age16.24, age25.34, age35.44, age45.54, age55.64, age65over)
+  # Perform data cleaning: keep only the columns that were simulated
+  cleaned_data() <- raw_data %>%
+  select(
+    date.mmm.yy., population_group, returned_from_housing, returned_to_shelter,
+    ageunder16, age16.24, age25.34, age35.44, age45.54, age55.64, age65over
+  )
 
 # Preview the cleaned data
 head(cleaned_data)

### 2024-09-27 scripts/03-test_data.R
styler scripts for test_data

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: test both simulated and cleaned data
 # Author: Xizi Sun
-# Date: 23 September 2024 
+# Date: 23 September 2024
 # Contact: xizi.sun@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Run the files "01-download_data.R" and "02-data_cleaning.R".
@@ -41,8 +41,8 @@ cat("Range of returned_to_shelter in simulated data:", range(simulated_data$retu
 cat("Range of returned_from_housing in simulated data:", range(simulated_data$returned_from_housing), "\n")
 
 # 5. Check for outliers in 'returned_to_shelter' and 'returned_from_housing'
-outliers_returned_to_shelter <- simulated_data$returned_to_shelter[simulated_data$returned_to_shelter > (mean(simulated_data$returned_to_shelter) + 3*sd(simulated_data$returned_to_shelter))]
-outliers_returned_from_housing <- simulated_data$returned_from_housing[simulated_data$returned_from_housing > (mean(simulated_data$returned_from_housing) + 3*sd(simulated_data$returned_from_housing))]
+outliers_returned_to_shelter <- simulated_data$returned_to_shelter[simulated_data$returned_to_shelter > (mean(simulated_data$returned_to_shelter) + 3 * sd(simulated_data$returned_to_shelter))]
+outliers_returned_from_housing <- simulated_data$returned_from_housing[simulated_data$returned_from_housing > (mean(simulated_data$returned_from_housing) + 3 * sd(simulated_data$returned_from_housing))]
 cat("Outliers in 'returned_to_shelter':", outliers_returned_to_shelter, "\n")
 cat("Outliers in 'returned_from_housing':", outliers_returned_from_housing, "\n")
 
@@ -84,8 +84,8 @@ cat("Range of returned_to_shelter in cleaned data:", range(cleaned_data$returned
 cat("Range of returned_from_housing in cleaned data:", range(cleaned_data$returned_from_housing), "\n")
 
 # 5. Check for outliers in 'returned_to_shelter' and 'returned_from_housing'
-outliers_returned_to_shelter_cleaned <- cleaned_data$returned_to_shelter[cleaned_data$returned_to_shelter > (mean(cleaned_data$returned_to_shelter) + 3*sd(cleaned_data$returned_to_shelter))]
[…]

### 2024-09-27 paper/paper.qmd
use styler to style the code and update paper.pdf

@@
   )
-
 ```
@@
 
-#Generated by ChatGPT
+# Generated by ChatGPT
 
@@
 
-#generated by ChatGPT
+# generated by ChatGPT
 
@@
 single_adult_population <- single_adult_population %>%
-  mutate(percentage = total_population / sum(total_population) * 100,
-         label = paste0(round(percentage, 1), "%"))  # Create a label with percentage
+  mutate(
+    percentage = total_population / sum(total_population) * 100,
+    label = paste0(round(percentage, 1), "%")
+  ) # Create a label with percentage
 
@@
 
-#developed by ChatGPT
+# developed by ChatGPT
 

### 2024-09-27 paper/paper.qmd
style the paper and fix typo

@@
 @sec-data-selection explains the reasons for not choosing another database.
-@sec-Results presents and describes the demographic distribution and trends of homelessness in graphs and charts.
-@sec-Discussion analyzes the information in the graphs in the context of the community and describes the limitations of the study.
+@sec-result presents and describes the demographic distribution and trends of homelessness in graphs and charts.
+@sec-discussion analyzes the information in the graphs in the context of the community and describes the limitations of the study.
 
@@
 
-# Discussion {#sec-data}
+# Discussion {#sec-discussion}
 

### 2024-09-27 paper/paper.qmd
use styler on the paper.qmd and update paper.pdf

@@
 
-# Calculate percentages
+# Calculate percentages and filter out zero percentages
 single_adult_population <- single_adult_population %>%
@@
     label = paste0(round(percentage, 1), "%")
-  ) # Create a label with percentage
+  ) %>%
+  filter(percentage > 0) # Exclude rows where percentage is 0
 
@@
   scale_fill_brewer(palette = "Set3") # Optional: Add color palette
+
 ```

### 2024-09-27 paper/paper.qmd
update reference and styler the paper

@@
 Visualizations were created using the ggplot2 package [@ggplot2], with data manipulation carried out via dplyr [@dplyr] and tidyr [@tidyr].
-Date handling was simplified with the lubridate package [@lubridate], and all R code was formatted using the styler package [@styler].
+Date handling was simplified with the lubridate package [@lubridate], and all R code was formatted using the styler package [@styler] and PDF file is generated by knitr [@knitr].
 Guidance on storytelling with data was drawn from Telling Stories with Data [@tellingstorieswithdata].
